% Generated from ICA_2.docx — re-run: python scripts/ica2_emit_latex.py
\chapter*{References}
\addcontentsline{toc}{chapter}{References}

\begin{enumerate}
  \item Ahmad, T. and Shah, M., 2024. Lightweight mini-batch federated learning technique for efficient attack detection in IoT systems. Journal of IoT Security / JISEM.
  \item Antonakakis, M., April, T., Bailey, M., Bernhard, M., Bursztein, E., Cochran, J., Durumeric, Z., Halderman, J.A., Invernizzi, L., Kallitsis, M. and Kumar, D., 2017. Understanding the Mirai botnet. In 26th USENIX Security Symposium, pp. 1093-1110.
  \item Bugcrowd, 2026. The Human Attack Surface: A Serious Threat to Cybersecurity. [online] Available at: https://www.bugcrowd.com/blog/the-human-attack-surface-a-serious-threat-to-cybersecurity/
  \item COMPSAC, 2025. A Lightweight Monitoring and Anomaly Detection Framework for IoT Devices. Proceedings of the IEEE Computer Society Signature Conference on Computers, Software and Applications.
  \item ENISA, 2023. Good Practices for IoT and Smart Infrastructures Tool. [online] European Union Agency for Cybersecurity. Available at: https://www.enisa.europa.eu/
  \item Fortinet, 2024. Why Are SMBs Most Vulnerable to Cyberattacks?. [online] Available at: https://www.fortinet.com/uk/resources/cyberglossary/smb-cyberattacks
  \item Future Internet, 2026. Multi-Attribute Physical-Layer Authentication Against Jamming and Battery-Depletion Attacks in LoRaWAN. Future Internet Journal, 18(38).
  \item Hassan, W. H., et al., 2020. Security Requirements for the Internet of Things: A Systematic Approach. National Institutes of Health (PMC).
  \item Information Commissioner's Office (ICO), 2024. Data protection and privacy for smart devices. [online] Available at: https://ico.org.uk/for-organisations/uk-gdpr-guidance-and-resources/
  \item IoT Analytics, 2024. State of IoT---Spring 2024. [online] IoT Analytics GmbH. Available at: https://iot-analytics.com/
  \item Journal of Engineering 360, 2024. Vulnerability of Wi-Fi wireless network to signal interference. [online] Available at: https://journalofengineering360.ajp.edu.pl/article/548861/en
  \item Kaspersky, 2024. Internet of Things security challenges and best practices. Kaspersky Resource Center.
  \item Keyfactor, 2024. IoT Device Security: Risks, Best Practices, and Protection Tips. Keyfactor Education Center.
  \item National Cyber Security Centre (NCSC), 2023. Smart devices in the home: a guide for businesses and individuals. [online] Available at: https://www.ncsc.gov.uk/guidance/smart-devices-in-the-home
  \item OAE Publish, 2023. Anatomy of attacks on IoT systems: review of attacks, impacts and countermeasures. Journal of Surveillance, Security and Safety.
  \item Open Worldwide Application Security Project (OWASP), 2024. OWASP Top 10 API Security Risks. [online] Available at: https://owasp.org/API-Security/
  \item R Discovery, 2025. Self-Adaptive Sensor Fault Detection in IoT Health Monitoring Using Federated Learning and Lightweight Transformers. [online] Available at: https://discovery.researcher.life/article/self-adaptive-sensor-fault-detection-in-iot-health-monitoring-using-federated-learning-and-lightweight-transformers/fee363dd342d3637957e563318e15e0e
  \item ResearchGate (Various Authors), 2025. Anomaly Detection in Real-Time IoT Systems Using Machine Learning. ResearchGate Publications.
  \item ResearchGate, 2025. An IoT-Based Smart Framework for a Human Heartbeat Rate Monitoring and Control System. [online] Available at: https://www.researchgate.net/publication/337275306\_An\_IoT-Based\_Smart\_Framework\_for\_a\_Human\_Heartbeat\_Rate\_Monitoring\_and\_Control\_System
  \item RiskRecon, 2023. Security in the Internet of Things. RiskRecon Security Blog.
  \item Roman, R., Zhou, J. and Lopez, J., 2013. On the features and challenges of security and privacy in distributed internet of things. Computer Networks, 57(10), pp. 2266-2279.
  \item SecuringSam, 2023. ``Hot Patching'' -- Seamlessly Protecting the Routers of Your SMB Customers. [online] Available at: https://securingsam.com/hot-patching-seamlessly-protecting-the-routers-of-your-smb-customers/
  \item Sicari, S., Rizzardi, A., Grieco, L.A. and Coen-Porisini, A., 2015. Security, privacy and trust in Internet of Things: The road ahead. Computer Networks, 76, pp. 146-164.
  \item SumatoSoft, 2025. Enterprise IoT Software Development for Businesses. [online] Available at: https://sumatosoft.com/services/enterprise-iot-development
  \item TechHive, 2024. Wi-Fi jamming attacks are on the rise. How they affect your smart home security. [online] Available at: https://www.techhive.com/article/2321439/wifi-jamming-attacks-smart-home-security.html
  \item Weber, R.H., 2010. Internet of Things -- New security and privacy challenges. Computer Law \& Security Review, 26(1), pp. 23-30.
  \item Zarpelão, B.B., Miani, R.S., Kawakani, C.T. and de Alvarenga, S.C., 2017. A survey of intrusion detection in Internet of Things. Journal of Network and Computer Applications, 84, pp. 25-37.
\end{enumerate}
